%HWFinal.tex
%
% Frank Sottile
% Thursday, 9 May 2002
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{amsart}
\usepackage{colordvi}
\usepackage{amssymb}


\headheight=8pt       \topmargin=-10pt
\topmargin=40pt     % LexmarkZ52 and Office printer
\textheight=760pt     \textwidth=456pt
\oddsidemargin=6pt   \evensidemargin=6pt

\DeclareMathOperator{\lcm}{lcm}

\begin{document}

\begin{center}
\Large\bf Final Homework\\
Math 411, Section 1\\
\large Frank Sottile\\
9 May 2002\vspace{10pt}
\end{center}

{\bf Problem Number 6 is now extra credit.  Due to a typographic error, 
     there are no solutions to the original problem, which has been fixed.}

This homework is due by the end of finals week (4PM Friday, 24 May), in the
envelope in Frank's mailbox in the 16th floor mathematics and statistics
departmental mail room, or under the door of
his office, LGRT 1230. 

This homework is worth 10\% of your final grade.
You must justify your answers for full credit.
Almost all of these problems are taken from the book.\bigskip

\begin{enumerate}

\item Let $SL(2,\mathbb{Z}_2)$ be the group of invertible $2\times 2$ matrices
      with entries from $\mathbb{Z}_2$, with operation matrix multiplication.
      (Ordinary matrix multiplication, but where we reduce the entries modulo 2.
       See Exercise 34 in Chapter 2 for the definition of this multiplication.)
     
      How many elements does $SL(2,\mathbb{Z}_2)$ have?
      Exhibit an isomorphism between $SL(2,\mathbb{Z}_2)$ and a familiar group.\medskip

\item Let $A$ be a (possibly infinite) abelian group.
      Prove that the set of elements of $A$ having finite order forms a
      subgroup of $A$. \medskip

\item Suppose that we have a finite group $G$ generated by two elements $a$ and 
      $b$ which satisfy  $a^4=b^2=e$, and $ba=a^3b$.
      Construct a Cayley table for $G$.
      Do these conditions uniquely determine $G$ up to isomorphism?\medskip

\item Let $G=\mathbb{Z}_3\oplus\mathbb{Z}_3\oplus\mathbb{Z}_3$, and let $H$ be
      the subgroup of $SL(3,\mathbb{Z}_3)$ consisting of upper triangular
      matrices with 1's on the diagonal
\[
     H\ =\ \left\{ \left[\begin{array}{ccc}1&a&b\\0&1&c\\0&0&1\end{array}
                   \right]\right\}\ .
\]
     Determine the number of elements of each order in both $G$ and $H$,
     and show that $G$ is {\it not} isomorphic to $H$.\medskip

\item Determine the number of elements and of subgroups of the following groups.
      $\mathbb{Z}_2$, $\mathbb{Z}_2\oplus\mathbb{Z}_2$, and
      $\mathbb{Z}_2\oplus\mathbb{Z}_2\oplus\mathbb{Z}_2$.
      How many subgroups do you expect $\bigl(\mathbb{Z}_2\bigr)^{\oplus n}$
      (the external direct sum of $n$ copies of $\mathbb{Z}_2$) to have?\medskip

\item Find three elements $\sigma$ of $S_9$ such that 
      $\sigma^3=(136)(297)(485)$.\medskip

\item Prove that a factor group of a cyclic group is cyclic.\medskip

\item Suppose that $G=H\times K$ and that $N$ is a normal subgroup of $H$.
      Prove that $N$ is a normal subgroup of $G$.\medskip

\item Find all the Sylow 3-subgroups of $A_4$, the alternating group
      of even permutations in $S_4$.\medskip

\item Let $G$ be a noncyclic group of order 21.
      How many Sylow 3-subgroups does $G$ have?
      How many elements of order 3 does $G$ have?
\end{enumerate}
\end{document}